Determinants of Soccer Players' Market Values in Major European Leagues

### 2024-11-27 paper/paper.qmd
Updated Draft

## Summary Statistics

Table 1 summarizes the descriptive statistics for the Premier League dataset.

@@

# Load data
england_data <- read_csv("../data/02-analysis_data/cleaned_england_data.csv")

# Generate summary statistics as a tidy tibble
summary_stats <- england_data %>%
  select(market_value, age, goals, assists, club_ranking, national_team_ranking, minutes_played) %>%
  summarise_all(list(mean = mean, sd = sd, min = min, max = max), na.rm = TRUE) %>%
  pivot_longer(cols = everything(),
               names_to = c("Variable", ".value"),
               names_sep = "_")

# Simplify and display the table using kable
kable(
  summary_stats,
  caption = "Summary Statistics for the Premier League Dataset",
  digits = 2
)
```

Key patterns across leagues:

- Average market value varies significantly, with the Premier League showing higher valuations.
- Goals and assists are more concentrated among forwards.
- Age distribution is consistent, skewed towards younger players.

@@

>>>>

## England

@@

# Load the dataset
england_data <- read_parquet("../data/02-analysis_data/cleaned_england_data.parquet")

# Scale down market value to millions
england_data <- england_data %>%
  mutate(market_value_million = market_value / 1e6)

# Individual plot for market value
england_market_value_plot <- ggplot(england_data, aes(x = market_value_million)) +
  geom_histogram(fill = "#D62728", color = "black", bins = 30) +
  theme_minimal() +
  labs(
    title = "Distribution of Market Value (in Million Euros)",
    x = "Market Value (Million Euros)",
    y = "Frequency"
  ) +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

england_market_value_plot
```

# Reshape data for faceting (excluding market_value)
data_long <- england_data %>%
  select(
    age,
    goals,
    assists,
    club_ranking,
    national_team_ranking,
    minutes_played
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Value"
  )

# Facet plot for the other variables
england_facet_plot <- ggplot(data_long, aes(x = Value)) +
  geom_histogram(fill = "#D62728", color = "black", bins = 30) +
  facet_wrap(~ Variable, nrow = 2, scales = "free") +
  theme_minimal() +
  labs(
    title = "Distribution of Other Variables",
    x = "Value",
    y = "Frequency"
  ) +
  theme(
    strip.text = element_text(size = 10, face = "bold"),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )
england_facet_plot
```

## France

# Load the dataset
france_data <- read_parquet("../data/02-analysis_data/cleaned_france_data.parquet")

# Scale down market value to millions
france_data <- france_data %>%
  mutate(market_value_million = market_value / 1e6)

# Individual plot for market value
france_market_value_plot <- ggplot(france_data, aes(x = market_value_million)) +
  geom_histogram(fill = "#1F77B4", color = "black", bins = 30) +
  theme_minimal() +
  labs(
    title = "Distribution of Market Value (in Million Euros)",
    x = "Market Value (Million Euros)",
    y = "Frequency"
  ) +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

france_market_value_plot
```

# Reshape data for faceting (excluding market_value)
data_long <- france_data %>%
  select(
    age,
    goals,
    assists,
    club_ranking,
    national_team_ranking,
    minutes_played
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Value"
  )

# Facet plot for the other variables
france_facet_plot <- ggplot(data_long, aes(x = Value)) +
  geom_histogram(fill = "#1F77B4", color = "black", bins = 30) +
  facet_wrap(~ Variable, nrow = 2, scales = "free") +
  theme_minimal() +
  labs(
    title = "Distribution of Other Variables",
    x = "Value",
    y = "Frequency"
  ) +
  theme(
    strip.text = element_text(size = 10, face = "bold"),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )
france_facet_plot
```

## Germany

# Load the dataset
germany_data <- read_parquet("../data/02-analysis_data/cleaned_germany_data.parquet")

# Scale down market value to millions
germany_data <- germany_data %>%
  mutate(market_value_million = market_value / 1e6)

# Individual plot for market value
germany_market_value_plot <- ggplot(germany_data, aes(x = market_value_million)) +
  geom_histogram(fill = "#000000", color = "white", bins = 30) +
  theme_minimal() +
  labs(
    title = "Distribution of Market Value (in Million Euros)",
    x = "Market Value (Million Euros)",
    y = "Frequency"
  ) +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

germany_market_value_plot
```

# Reshape data for faceting (excluding market_value)
data_long <- germany_data %>%
  select(
    age,
    goals,
    assists,
    club_ranking,
    national_team_ranking,
    minutes_played
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Value"
  )

# Facet plot for the other variables
germany_facet_plot <- ggplot(data_long, aes(x = Value)) +
  geom_histogram(fill = "#000000", color = "white", bins = 30) +
  facet_wrap(~ Variable, nrow = 2, scales = "free") +
  theme_minimal() +
  labs(
    title = "Distribution of Other Variables",
    x = "Value",
    y = "Frequency"
  ) +
  theme(
    strip.text = element_text(size = 10, face = "bold"),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )
germany_facet_plot
```

## Italy

# Load the dataset
italy_data <- read_parquet("../data/02-analysis_data/cleaned_italy_data.parquet")

# Scale down market value to millions
italy_data <- italy_data %>%
  mutate(market_value_million = market_value / 1e6)

# Individual plot for market value
italy_market_value_plot <- ggplot(italy_data, aes(x = market_value_million)) +
  geom_histogram(fill = "#2CA02C", color = "black", bins = 30) +
  theme_minimal() +
  labs(
    title = "Distribution of Market Value (in Million Euros)",
    x = "Market Value (Million Euros)",
    y = "Frequency"
  ) +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

italy_market_value_plot
```

# Reshape data for faceting (excluding market_value)
data_long <- italy_data %>%
  select(
    age,
    goals,
    assists,
    club_ranking,
    national_team_ranking,
    minutes_played
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Value"
  )

# Facet plot for the other variables
italy_facet_plot <- ggplot(data_long, aes(x = Value)) +
  geom_histogram(fill = "#2CA02C", color = "black", bins = 30) +
  facet_wrap(~ Variable, nrow = 2, scales = "free") +
  theme_minimal() +
  labs(
    title = "Distribution of Other Variables",
    x = "Value",
    y = "Frequency"
  ) +
  theme(
    strip.text = element_text(size = 10, face = "bold"),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )
italy_facet_plot
```

## Spain

# Load the dataset
spain_data <- read_parquet("../data/02-analysis_data/cleaned_spain_data.parquet")

# Scale down market value to millions
spain_data <- spain_data %>%
  mutate(market_value_million = market_value / 1e6)

# Individual plot for market value
spain_market_value_plot <- ggplot(spain_data, aes(x = market_value_million)) +
  geom_histogram(fill = "#FFB000", color = "black", bins = 30) +
  theme_minimal() +
  labs(
    title = "Distribution of Market Value (in Million Euros)",
    x = "Market Value (Million Euros)",
    y = "Frequency"
  ) +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

spain_market_value_plot
```

# Reshape data for faceting (excluding market_value)
data_long <- spain_data %>%
  select(
    age,
    goals,
    assists,
    club_ranking,
    national_team_ranking,
    minutes_played
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Variable",
    values_to = "Value"
  )

# Facet plot for the other variables
spain_facet_plot <- ggplot(data_long, aes(x = Value)) +
  geom_histogram(fill = "#FFB000", color = "black", bins = 30) +
  facet_wrap(~ Variable, nrow = 2, scales = "free") +
  theme_minimal() +
  labs(
    title = "Distribution of Other Variables",
    x = "Value",
    y = "Frequency"
  ) +
  theme(
    strip.text = element_text(size = 10, face = "bold"),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )
spain_facet_plot
```

@@

# Pivot data to create comparison tables
comparison_table <- final_results %>%
  select(term, estimate, League) %>%
  pivot_wider(names_from = League, values_from = estimate)

# Intercept comparison table
intercept_table <- comparison_table %>%
  filter(term == "(Intercept)") %>%
  select(-term)

# Display Intercept Table
kable(
  intercept_table,
  digits = 3,
  caption = "Intercept Estimates Across Leagues"
)
```

# Age comparison table
age_table <- comparison_table %>%
  filter(term == "age") %>%
  select(-term)

# Display Age Table
kable(
  age_table,
  digits = 3,
  caption = "Age Coefficient Estimates Across Leagues"
)
```

# Goals comparison table
goals_table <- comparison_table %>%
  filter(term == "goals") %>%
  select(-term)

# Display Goals Table
kable(
  goals_table,
  digits = 3,
  caption = "Goals Coefficient Estimates Across Leagues"
)
```

# Assists comparison table
assists_table <- comparison_table %>%
  filter(term == "assists") %>%
  select(-term)

# Display Assists Table
kable(
  assists_table,
  digits = 3,
  caption = "Assists Coefficient Estimates Across Leagues"
)
```

# Club Ranking comparison table
club_ranking_table <- comparison_table %>%
  filter(term == "club_ranking") %>%
  select(-term)

# Display Club Ranking Table
kable(
  club_ranking_table,
  digits = 3,
  caption = "Club Ranking Coefficient Estimates Across Leagues"
)
```

# Appendix {#sec-appendix}

## Dataset and Graph Sketches

## Model Results
